---
title: "Hands On Exercise 2"
author: "Seng Jing Yi"
date: "April 19, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
format: html
editor: visual
---

## 2.1 Objectives

In this chapter, you will be introduced to several ggplot2 extensions for creating more elegant and effective statistical graphics. By the end of this exercise, you will be able to:

-   control the placement of annotation on a graph by using functions provided in ggrepel package,

-   create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages,

-   plot composite figure by combining ggplot2 graphs by using patchwork package.

## 2.2 Getting Started

### 2.2.1 Installing and loading required libraries

-   ggrepel: an R package provides geoms for ggplot2 to repel overlapping t ext labels.

-   ggthemes: an R package provides some extra themes, geoms, and scales for ‘ggplot2’.

-   hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.

-   patchwork: an R package for preparing composite figure created using ggplot2.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, readr) 
```

### 2.2.2 Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## **2.3 Beyond ggplot2 Annotation: ggrepel**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### 2.3.1. **Working with ggrepel**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.4 Beyond ggplot 2 Themes

ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html), they are: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, and `theme_void()`.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### **2.4.1 Working with ggtheme package**

[**ggthemes**](https://cran.r-project.org/web/packages/ggthemes/) provides [‘ggplot2’ themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) that replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), ‘Stata’, ‘Excel’, and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### 2.4.2 Working with hrbthems package

[**hrbrthemes**](https://cinc.rud.is/web/packages/hrbrthemes/) package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

## 2.5 **Beyond Single Graph**

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs. In this section, you will learn how to create composite plot by combining multiple graphs. First, let us create three statistical graphics by using the code chunk below.

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p1
```

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p2
```

Scatterplot for English Score vs Math Score:

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

p3
```

### **2.5.1 Creating Composite Graphics: pathwork methods**

There are several ggplot2 extension’s functions support the needs to prepare composite figure by combining several graphs such as [`grid.arrange()`](https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html) of **gridExtra** package and [`plot_grid()`](https://wilkelab.org/cowplot/reference/plot_grid.html) of [**cowplot**](https://wilkelab.org/cowplot/) package. In this section, I am going to shared with you an ggplot2 extension called [**patchwork**](https://patchwork.data-imaginist.com/) which is specially designed for combining separate ggplot2 graphs into a single figure.

Patchwork package has a very simple syntax where we can create layouts super easily. Here’s the general syntax that combines:

-   Two-Column Layout using the Plus Sign +.

-   Parenthesis () to create a subplot group.

-   Two-Row Layout using the Division Sign `/`

### **2.5.2 Combining two ggplot2 graphs**

Figure in the tabset below shows a composite of two histograms created using patchwork. Note how simple the syntax used to create the plot!

```{r}
p1 + p2
```

### **2.5.3 Combining three ggplot2 graphs**

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

-   “/” operator to stack two ggplot2 graphs,

-   “\|” operator to place the plots beside each other,

-   “()” operator the define the sequence of the plotting.

```{r}
(p1 / p2) | p3
```

### **2.5.4 Creating a composite figure with tag**

In order to identify subplots in text, **patchwork** also provides auto-tagging capabilities as shown in the figure below.

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### **2.5.5 Creating figure with insert**

Beside providing functions to place plots next to each other based on the provided layout. With [`inset_element()`](https://patchwork.data-imaginist.com/reference/inset_element.html) of **patchwork**, we can place one or several plots or graphic elements freely on top or below another plot.

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### **2.5.6 Creating a composite figure by using patchwork and ggtheme**

Figure below is created by combining patchwork and theme_economist() of ggthemes package discussed earlier.

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
