---
title: "Take-home Exercise 3"
author: "Seng Jing Yi"
date: "May 18, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
editor: visual

format:
  html:
    code-fold: true
---

# Exploring Vast Challenge 2024 - Mini-challenge 2

## 1. Introduction:

### 1.1 Setting the scene:

FishEye has learned that SouthSeafood Express Corp has been caught fishing illegally. The scandal caused a major disruption in the close-knit fishing community. FishEye has been collecting data on ship movements and shipping records in hopes that they could assemble a cohesive store of knowledge that will allow them to better understand local commercial fishing behavior.

### 1.2 Key Objectives of this assignment:

FishEye analysts need your help to perform geographic and temporal analysis of the CatchNet data so they can prevent illegal fishing from happening again. This includes:

1.  Visualization system to associate vessels with their probable cargos, including seasonality analysis to detect any anomalies in port exit records.
2.  Visualization to illustrate inappropriate behavior of SouthSeafood Express Corp Vessels, in terms of vessel movements, catch contents, regions where illegal fishing is performed, telling-signs of suspicious behaviors.
3.  Identifying vessels engaging in similar behavious to SouthSeafood Express Corp.
4.  Visualizing changes in fishing activities after SouthSeafood Express was caught.

## 2. Data Import & Processing

### 2.1 Loading relevant packages and data

```{r}
# load package
pacman::p_load(jsonlite, tidyverse, tidyr, 
               knitr, lubridate, dplyr, 
               igraph, ggraph, ggdist, ggplot2, 
               SmartEDA, sf, tidygraph, reshape2, readr,
               DT, patchwork,plotly)

# load data
mc2_data <- fromJSON("data/mc2.json")
oceanus_geog = st_read("data/Oceanus Geography.geojson")
```

### 2.2. Data Cleaning

#### 2.2.1 Extracting edges & understanding edge tibble data table.

1.  Extracting JSON file `edges` to tibble data frame and removing duplicates
2.  Converting to the correct data types for datetime
3.  Renaming columns starting with "\_ xx\_" to minimize downstream syntax errors.
4.  Creating subsets of data tables based on column of "type"
5.  Renaming of columns to provide context and unique identifiers for downstream data mapping.
6.  Resultant data sets include:

+------------------------------------+--------------------------------------------------------------------------------------------------------+
| Resultant Edges Data Set           | Unique columns                                                                                         |
+====================================+========================================================================================================+
| Transactions (i.e., tx_c)          | Cargo_id, Destination, Transaction date, Fish species                                                  |
+------------------------------------+--------------------------------------------------------------------------------------------------------+
| Transponder Ping (i.e., E_Tping_c) | Start_time, Dwell, Ping source, Vessel_id                                                              |
+------------------------------------+--------------------------------------------------------------------------------------------------------+
| Habor Arrival Report (E_Hbrpt_c)   | Vessel_id, Port, Key, Arrival date, Port master, Aphorism, Holiday Greeting, Saying of the Sea, Wisdom |
+------------------------------------+--------------------------------------------------------------------------------------------------------+

```{r}
#assigning to mc2_edges2
mc2_edges <- as_tibble(mc2_data$links) %>% 
  distinct() 

#correcting data type - converting to date format
mc2_edges$time <- as_datetime(mc2_edges$time)
mc2_edges$"_last_edited_date" <- as_datetime(mc2_edges$"_last_edited_date")
mc2_edges$"_date_added" <- as_datetime(mc2_edges$"_date_added")
mc2_edges$date <- as_datetime(mc2_edges$date)


#renaming headers with "_" to prevent errors. 
mc2_edges <- mc2_edges %>%
  rename("last_edited_by" = "_last_edited_by",
         "date_added" = "_date_added",
         "last_edited_date" = "_last_edited_date",
         "raw_source" = "_raw_source",
         "algorithm" = "_algorithm") 

#checking data format
#glimpse(mc2_edges)

# Breaking into subsets based on event category
E_TransponderPing <- subset(mc2_edges,  mc2_edges$type == "Event.TransportEvent.TransponderPing")
E_HarborRpt <- subset(mc2_edges,  mc2_edges$type == "Event.HarborReport")
E_Tx <- subset(mc2_edges, mc2_edges$type == "Event.Transaction")

# Dropping columns that are NULL and renaming variables to separate them

# Transactions
E_Tx_c <- E_Tx %>%
  rename(
  cargo_id = source, 
  dest = target,
  tx_date = date) %>%
  select(-c(key, algorithm, `raw_source`, `type`, `data_author`, `aphorism`, `holiday_greeting`, `wisdom`, `saying of the sea`, `time`, `dwell`)) 

# Separating the fish species for the respective cargo 
tx_sub1 <- E_Tx_c[grep("^City of", E_Tx_c$dest), ]
tx_sub2 <- E_Tx_c[!grepl("^City of", E_Tx_c$dest), ]
tx_sub2 <- tx_sub2 %>% rename(fish_species = dest)

tx_c <- left_join(tx_sub1, tx_sub2 %>% select(cargo_id, fish_species), by = "cargo_id")

# Dropping raw source: All Oceanus Centralized Export/Import Archive and Notatification Service (OCEANS)
# Dropping algorithm: CatchMate ('arrrr' edition)
# Null columns - Data author, aphorism, holiday_greeting, wisdom, saying of the sea, time, dwell

# Transponder Ping
E_Tping_c <- E_TransponderPing %>%
  rename(vessel_id = target, 
         ping_source = source,
         start_time = time) %>%
  select(-c(key, `algorithm`, `raw_source`, `type`, `date`, `data_author`, `aphorism`, `holiday_greeting`, `wisdom`, `saying of the sea`)) 

#Dropping raw_source: All Oceanus Vessel Locator System
#Dropping algorithm: All OVLS-Catch&Hook
# Null columns - Date, Data author, aphorism, holiday_greeting, wisdom, saying of the sea

# Habour Report
E_Hbrpt_c <- E_HarborRpt %>% rename(
  vessel_id = source, 
  port = target, 
  arr_date = date, 
  port_master = data_author, 
  saying = `saying of the sea`
) %>%
  select(-c(`algorithm`, `type`, `time`, `dwell`)) 

#Dropping algorithm: All HarborReportMaster 3.11
#Retain raw_source: Differing values depending on which Port / City
# check meaning of key (0,1,2)

```

#### 2.2.2 Extracting nodes & understanding nodes tibble data table.

Repeating similar steps for `nodes` records, we obtained the following resultant data set. To minimize the number of data sets, we appended nodes information on vessels and region into a combined data set, with an assigned label to identify it's category. (e.g., vessel_type)

+------------------------------------------+--------------------------------------------------------------------------------------------+
| Resultant Nodes Data Sets                | Unique Columns                                                                             |
+==========================================+============================================================================================+
| N_fish (Fishes and their description)    | Fish_id, fish species                                                                      |
+------------------------------------------+--------------------------------------------------------------------------------------------+
| N_Delivery_Doc (Cargo and details)       | Cargo_id, quantity in tons, delivery date                                                  |
+------------------------------------------+--------------------------------------------------------------------------------------------+
| N_vessel (Vessels and their description) | Vessel_id, vessel name, vessel company, flag country, tonnage, overall length, vessel type |
+------------------------------------------+--------------------------------------------------------------------------------------------+
| Location_legend (Point, City, Region)    | Area, activities, kind, fish_species                                                       |
+------------------------------------------+--------------------------------------------------------------------------------------------+

```{r}
#segmenting nodes data and checking for distinct records
mc2_nodes <- as_tibble(mc2_data$nodes) %>%
  distinct()

#renaming to remove the "_" 
mc2_nodes <- mc2_nodes %>%
  rename("last_edited_by" = "_last_edited_by",
         "date_added" = "_date_added",
         "last_edited_date" = "_last_edited_date",
         "raw_source" = "_raw_source",
         "algorithm" = "_algorithm") 

#tidying the text data to remove nested list.

mc2_nodes_tidied <- mc2_nodes %>%
  mutate(Activities = gsub("c[(]", "", Activities)) %>% 
  mutate(Activities = gsub("\"", "", Activities)) %>%
  mutate(Activities = gsub("[)]", "", Activities)) 

mc2_nodes_tidied <- mc2_nodes_tidied %>%
  mutate(fish_species_present = gsub("c[(]", "", fish_species_present)) %>% 
  mutate(fish_species_present = gsub("\"", "", fish_species_present)) %>%
  mutate(fish_species_present = gsub("[)]", "", fish_species_present)) 

# Creating subset on nodes information
N_fish <- subset(mc2_nodes_tidied,  mc2_nodes_tidied$type == "Entity.Commodity.Fish") %>%
  select_if(~ !any(is.na(.))) %>% 
  select(-c(`type`, `raw_source`, `algorithm`, `Activities`, `fish_species_present`)) %>%
  rename(fish_species = name, 
         fish_id = id)

NL_City <- subset(mc2_nodes_tidied,  mc2_nodes_tidied$type == "Entity.Location.City") %>%
  select_if(~ !any(is.na(.))) %>%
  select(-c(`raw_source`, `algorithm`, `type`, `fish_species_present`)) %>%
  rename(city_name = Name, 
         city_id = id)


NL_Point <- subset(mc2_nodes_tidied,  mc2_nodes_tidied$type == "Entity.Location.Point") %>%
  select_if(~ !any(is.na(.))) %>%
  select(-c(`raw_source`, `algorithm`, `kind`, `fish_species_present`)) %>%
  rename(point_name = Name, 
         point_id = id)

## Need to tidy NL Region
NL_Region <- subset(mc2_nodes_tidied,  mc2_nodes_tidied$type == "Entity.Location.Region") %>%
  select_if(~ !any(is.na(.))) %>%
  select(-c(`raw_source`, `algorithm`, `type`, `Description`)) %>%
  rename(region_name = Name, 
         region_id = id, 
         region_kind = kind)

N_Delivery_doc <- subset(mc2_nodes_tidied,  mc2_nodes_tidied$type == "Entity.Document.DeliveryReport") %>%
  select_if(~ !any(is.na(.))) %>%
  rename(deliver_date = date,
         cargo_id = id) %>%
  select(-c(`algorithm`, `type`, `raw_source`, `Activities`, `fish_species_present`)) 

## consider adding back more columns, it dropped some columns where values were partial NA

N_vessel <- mc2_nodes_tidied %>%
  filter(grepl("Entity.Vessel", type)) %>%
  mutate(vessel_type = case_when(
    grepl("FishingVessel", type, ignore.case = TRUE) ~ "Fishing",
    grepl("Ferry.Passenger", type, ignore.case = TRUE) ~ "Ferry_Passenger",
    grepl("Ferry.Cargo", type, ignore.case = TRUE) ~ "Ferry_Cargo",
    grepl("Research", type, ignore.case = TRUE) ~ "Research", 
    grepl("Other", type, ignore.case = TRUE) ~ "Other", 
    grepl("Tour", type, ignore.case = TRUE) ~ "Tour", 
    grepl("CargoVessel", type, ignore.case = TRUE) ~ "Cargo_Vessel"
    )) %>% 
  mutate(company = ifelse(is.na(company), "Unknown", company)) %>% # Handle NA values by replacing NA with unknown
  select(-c(`algorithm`, `type`, `raw_source`, `Activities`, `fish_species_present`, `Description`, `kind`, `style`, `name`, `qty_tons`,`date`)) %>%
  rename(vessel_id = id, 
         vessel_name = Name, 
         vessel_company = company)


# Further exploring records where there is null values
partial_na_records <- N_vessel[!complete.cases(N_vessel), ]
kable(partial_na_records)

# merging ping with location details (region_id, city_id, point_id)
## See how to deal with list within activites, and how to include region_kind
## See if want to retain some description to identify point, region vs city

city_legend <- NL_City %>% 
  select(c(`city_id`, `Activities`, `kind`)) %>%
  mutate(fish_species_present = "NA") %>%
  rename(area = city_id)

point_legend <- NL_Point %>% 
  select(c(`point_id`, `Activities`))  %>%
  mutate(kind = "point", 
         fish_species_present = "NA") %>%
  rename(area = point_id)

region_legend <- NL_Region %>% 
  select(c(`region_id`, `Activities`, `region_kind`, `fish_species_present`))  %>%
  rename(area = region_id, kind = region_kind)

location_legend <- rbind(city_legend, point_legend, region_legend) 

```

#### 2.2.5 Merging back the data after processing.

To incorporate context of the nodes details into the various edges, the related description of the nodes were appended to edge data sets. This helped to streamline the records into 3 consolidated data sets.

+-------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Consolidated Data Set                                                                     | Resultant Edge Data Set                                                                                             | Mapped with Nodes Data Set                                                                             |
+===========================================================================================+=====================================================================================================================+========================================================================================================+
| Transaction (with cargo weight)                                                           | `tx_c`: Cargo_id, Destination, Transaction date, Fish species                                                       | `N_delivery_doc`: Cargo_id, quantity in tons, delivery date                                            |
+-------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Ping activity (with vessel details by vessel_id, possible fish caught by location_legend) | `E_Tping_c`: Start_time, Dwell, Ping source, Vessel_id                                                              | `N_vessel`: Vessel_id, vessel name, vessel company, flag country, tonnage, overall length, vessel type |
|                                                                                           |                                                                                                                     |                                                                                                        |
|                                                                                           |                                                                                                                     | `location_legend`: Area, activities, kind, fish_species                                                |
+-------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------+
| Habor Arrival Report\                                                                     | `E_Hbrpt_c`: Vessel_id, Port, Key, Arrival date, Port master, Aphorism, Holiday Greeting, Saying of the Sea, Wisdom | `N_vessel`: Vessel_id, vessel name, vessel company, flag country, tonnage, overall length, vessel type |
| (with vessel details)                                                                     |                                                                                                                     |                                                                                                        |
+-------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------+

```{r}
# merging the quantity from delivery_doc with transaction based on cargo id
tx_qty <- left_join(tx_c, N_Delivery_doc %>% select(cargo_id, qty_tons, deliver_date), by = "cargo_id")

# Merging habor report with vessel details to identify type of vessel, tonnage and overall length 

E_Hbrpt_v <- left_join(E_Hbrpt_c, N_vessel %>% select(vessel_id, vessel_name, tonnage, length_overall, flag_country, vessel_company, vessel_type), by = "vessel_id")

# Merging ping transaction with the location type
E_Tping_c2 <- E_Tping_c %>% rename(area = ping_source)
ping_activity <- left_join(E_Tping_c2, location_legend %>% select(area, Activities, kind, fish_species_present), by = "area")

#write_csv(ping_activity, "movement.csv")
# suggest drop all date added, last edits and edited by - used for biasness analysis, not so much for mc2

```

## 3. Exploratory Data Analysis

#### 3.0 Understanding areas in Oceanus

```{r}
ggplot(data = oceanus_geog) + geom_sf()
```

#### 3.1 Understanding possible fish species and signs of illegal fishing

::: callout-note
## Insights

-   3 fish species are only present in ecological reserves: (1) Offidiaa/Piscis osseus, (2) Sockfish/Pisces foetida and (3) Helenaa/Pisces satis. Hence, any cargo with these fish species have likely violated fishing regulations and fished in ecological reserves. (1) Don Limpet Preserve, (2) Ghoti Preserve and (3) Nemo Reefs.
:::

```{r}
## Check if code can still be run given above amendment of string split

# Formatting region data to identify fish type in region
region_species <- NL_Region %>%
  mutate(fish_species_present = gsub('c\\(|\\)|"', "", fish_species_present), 
    fish_species_present = strsplit(as.character(fish_species_present), ", ")) %>%
  unnest(fish_species_present) %>%
  mutate(presence = 1) %>%
  spread(key = fish_species_present, value = presence, fill = 0)

region_species_c <- region_species %>% 
  select( -region_id, -last_edited_by, -last_edited_date, -date_added)

kable(region_species_c)

#Aligning the naming convention for fish species

region_species_id <- region_species %>% rename(
  gadusnspecificatae4ba = `Cod/Gadus n.specificatae`, 
  piscesfrigus900 = `Birdseye/Pisces frigus`, 
  piscesfoetidaae7 = `Sockfish/Pisces foetida`, #illegal
  labridaenrefert9be = `Wrasse/Labridae n.refert`, 
  habeaspisces4eb = `Beauvoir/Habeas pisces`, 
  piscissapidum9b7 = `Harland/Piscis sapidum`, 
  thunnininveradb7 = `Tuna/Thunnini n.vera`, 
  piscisosseusb6d = `Offidiaa/Piscis osseus`, #illegal
  piscessatisb87 = `Helenaa/Pisces satis`) #illegal

# write_csv(region_species, "species_region_id.csv")

# Location of Salmon not identified (Extra value of Salmon in Fish node)
# Check what to do with cargos with salmon - oncorhynchusrosea790

# identifying occurence of illegal fish species at various port - identifying all unique fish species in cargo report (tx_qty)

unique_fish_cargo <- unique(tx_qty$fish_species)
unique_fish_cargo

## assign specific colors to fish species, red for illegal. 

fish_species_color <- c("piscesfoetidaae7" = "#FF6666", 
                        "piscisosseusb6d" = "#FF9999", 
                        "piscessatisb87" = "#FFCCCC", 
                        "gadusnspecificatae4ba" = "#CCE5FF", 
                        "piscissapidum9b7" = "#99CCFF", 
                        "habeaspisces4eb" = "#66B2ff", 
                        "piscesfrigus900" = "#CCE5FF", 
                        "oncorhynchusrosea790" = "#FFFF99", 
                        "labridaenrefert9be" = "#99CCFF", 
                        "thunnininveradb7" = "#66b2ff"
                        )

# Summarise total tons of fish per location
total_qty_tons_per_dest <- tx_qty %>%
  group_by(dest, fish_species) %>%
  summarize(total_qty_tons = sum(qty_tons, na.rm = TRUE)) %>%
  ungroup()

# Reordering levels for fish species for tidier plot
total_qty_tons_per_dest$fish_species <- factor(
  total_qty_tons_per_dest$fish_species, 
  levels = c("piscesfoetidaae7","piscisosseusb6d", "piscessatisb87", #illegal
             "oncorhynchusrosea790", # unidentified - Salmon
             "gadusnspecificatae4ba", "piscissapidum9b7", "habeaspisces4eb", 
             "piscesfrigus900", "labridaenrefert9be", "thunnininveradb7"))

# Plot
ggplot(total_qty_tons_per_dest, aes(x = dest, y = total_qty_tons, fill = fish_species)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fish_species_color) +
  labs(title = "Total Quantity of Fish by Destination and Species",
       x = "Destination",
       y = "Total Quantity (tons)",
       fill = "Fish Species") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### 3.2. Nodes on Vessels - Understanding network of vessels in relation to companies

-   Comparables to South Seafood Express Corp

#### Exploring vessel tonnage by type and company

```{r}
scatter_ton_len <- ggplot(data= N_vessel, 
            aes(x= tonnage, 
                y= length_overall, 
                color= vessel_type)) +
  geom_point() + 
  theme(legend.position = "top")

vessel_count <- N_vessel %>%
  group_by(vessel_type) %>%
  summarize(vessel_no = n()) %>% 
  mutate(vessel_type = reorder(vessel_type, - vessel_no))


bar_vessel_type <- ggplot(data = vessel_count, 
                          aes(x = vessel_type, 
                              y = vessel_no)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = vessel_no), vjust = -0.8, size = 2) + 
  theme(axis.text.x = element_text(size = 5))

scatter_ton_len | bar_vessel_type
```

Focusing on Fishing vessels and Cargo vessels, identifying the distribution of ships and companies owning this ship

```{r}

fishing_v <- N_vessel %>% filter(vessel_type == "Fishing")
cargo_v <- N_vessel %>% filter(vessel_type == "Cargo_Vessel")

fishing_v_dist <- ggplot(fishing_v, aes(x = tonnage)) +
  geom_histogram(binwidth = 500, fill = "blue", color = "black") +
  labs(title = "Distribution of Fishing Vessels by Tonnage",
       x = "Tonnage",
       y = "Frequency") +
  theme_minimal() + coord_flip()

# include boxplot to see quartile 

cargo_v_dist <- ggplot(cargo_v, aes(x = tonnage)) +
  geom_histogram(binwidth = 500, fill = "orange", color = "black") +
  labs(title = "Distribution of Cargo Vessels by Tonnage",
       x = "Tonnage",
       y = "Frequency") +
  theme_minimal() + 
  coord_flip()

#summary(fishing_v) - min: 100, q1: 600, median: 2400, q3: 4850, max: 17200 - investigate
#summary(cargo_v) - min: 2100, q1: 2100, median: 23750, q3: 74925, max: 76300 - Investigate 

fishing_v_dist | cargo_v_dist
```

Insights derived:

1.  Exception size of fishing vessels (Vessels \> 12,500 tons). Possible explanation: Fishing of different nature - to search further on types of fishing vessels.

```{r}
# understanding which vessel 
abn_fish_vessel <- fishing_v %>% filter(tonnage > 12500)

sel_abn_fish_vessel <- abn_fish_vessel %>% 
  select(vessel_name, flag_country, vessel_company) 

kable(sel_abn_fish_vessel)

# understanding region that vessel spends time at 

abn_fish_v_activity <- ping_activity %>% 
  filter(vessel_id %in% abn_fish_vessel$vessel_id) 
  
abn_fish_v_activity_sum <- abn_fish_v_activity %>%
  group_by(vessel_id, area) %>%
  summarise(median_dwell = median(dwell, na.rm = TRUE))

# Contrasting with the median time spent by fishing vessels 

fish_v_activity <- ping_activity %>% 
  filter(vessel_id %in% fishing_v$vessel_id) 
  
fish_v_activity_sum <- fish_v_activity %>%
  group_by(vessel_id, area) %>%
  summarise(median_dwell = median(dwell, na.rm = TRUE))

# kable(ETping_vessel)
# Flag out vessels that entered restricted zone

fishing_dwell <- ggplot(fish_v_activity, aes(x = area, y = dwell)) +
  geom_boxplot() +
  geom_point(data = subset(fish_v_activity, vessel_id == "marinemarauder8c9"),
             aes(x = area, y = dwell), color = "red", size = 1) +
  geom_point(data = subset(fish_v_activity, vessel_id == "pikepirate89a"),
             aes(x = area, y = dwell), color = "blue", size = 1) +
  theme_minimal() +
  labs(title = "Median Dwell Time by Area", 
       x = "Area", 
       y = "Median Dwell Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 600000)) + # limit dwell 
  coord_flip()
```

2.  Wide variation of size for cargo vessels ( Grp 1: \< 5000 tons, Grp 2: 75,000 \> 5000 tons)

    Observed that vessel company is unknown for all 39 records in group 1, and all unknown for all 33 records in group 2

```{r}

abn_cargo_vessel_s <- cargo_v %>% filter(tonnage < 5000)
abn_cargo_vessel_l <- cargo_v %>% filter(tonnage > 70000)

sel_abn_cargo_vessel_s <- abn_cargo_vessel_s %>% 
  select(vessel_name, flag_country, vessel_company) 

sel_abn_cargo_vessel_l <- abn_cargo_vessel_l %>% 
  select(vessel_name, flag_country, vessel_company) 

# display records 
kable(sel_abn_cargo_vessel_s)
kable(sel_abn_cargo_vessel_l)

# subset cargo vessel activity
cargo_v_activity <- ping_activity %>% 
  filter(vessel_id %in% cargo_v$vessel_id) 

# extracting list of vessel id for large and small cargo
selected_vessel <- abn_cargo_vessel_s$vessel_id
selected_vessel2 <- abn_cargo_vessel_l$vessel_id

# plotting
cargo_dwell <- ggplot(cargo_v_activity, aes(x = area, y = dwell)) +
  geom_boxplot() +
  geom_point(data = subset(cargo_v_activity, vessel_id == selected_vessel),
             aes(x = area, y = dwell), color = "orange", size = 1) +
  geom_point(data = subset(cargo_v_activity, vessel_id == selected_vessel2),
             aes(x = area, y = dwell), color = "purple", size = 1) +
  theme_minimal() +
  labs(title = "Median Dwell Time by Area", 
       x = "Area", 
       y = "Median Dwell Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_y_continuous( limits = c(0, 20000)) + #zoom into 2 * 10^5
  coord_flip()

fishing_dwell | cargo_dwell
```

Exploring ships from "Unknown companies" - Unregistered ships - Transhipment cargo vessel, only 1 registered with Oceanus, else others by nations.

```{r}
# exploring ships that have unknown companies

unknown_v <- N_vessel %>% filter(vessel_company == "Unknown")

# Summarizing to see where unknown vessels come from
unknown_v_sum <- unknown_v %>% 
  group_by(vessel_type, flag_country) %>% 
  summarize(vessel_count = n()) %>%
  mutate(flag_country = reorder(flag_country, - vessel_count))

unknown_dist <- ggplot(unknown_v_sum, 
                       aes(x = flag_country, y = vessel_count, 
                           fill = vessel_type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Vessel Count by Flag Country and Vessel Type",
       x = "Flag Country",
       y = "Vessel Count",
       fill = "Vessel Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5)) 

unknown_dist 

kable(unknown_v_sum)

# Identifying regions that unknown vessels are at 
unknown_v_list <- unknown_v$vessel_id

vessel_dwell <- ggplot(ping_activity, aes(x = area, y = dwell)) +
  geom_boxplot() +
  geom_point(data = subset(ping_activity, vessel_id == unknown_v_list),
             aes(x = area, y = dwell), color = "hotpink", size = 1) +
  theme_minimal() +
  labs(title = "Median Dwell Time by Area", 
       x = "Area", 
       y = "Median Dwell Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_y_continuous( limits = c(0, 500000)) + #zoom into 2 * 10^5
  coord_flip()

vessel_dwell

# identifying outliers - nemo-reef > 2,000,000
# identifying ships that are way above median and "unknown"
```

**Exploring the cargo records**

-   Consistent record where qty_tons \<=0 across all 5 cities

-   However, variation in fish species where qty_tons \<= 0 for cargo

```{r}
#exploring abnormal records where qty_tons of cargo delivered = 0 or is negative

ggplot(tx_qty, aes(x = fish_species, y = qty_tons)) + 
  geom_boxplot(
    width = .15, 
    outlier.color = NA 
  ) +
  ggdist::stat_dots(data = subset(tx_qty, qty_tons > 0),
    side = "right", 
    justification = -0.2,
    binwidth = 1, 
    dotsize = 0.15) + 
   ggdist::stat_dots(
    data = subset(tx_qty, qty_tons <= 0),
    side = "right",
    justification = -0.2,
    binwidth = 1,
    dotsize = 0.15,
    color = "red") + 
  geom_hline(yintercept = 0, color = "blue") + 
  coord_flip() 
  
```

```{r}
#exploring abnormal records where qty_tons of cargo delivered = 0 or is negative

ggplot(tx_qty, aes(x = dest, y = qty_tons)) + 
  geom_boxplot(
    width = .15, 
    outlier.color = NA 
  ) +
  ggdist::stat_dots(data = subset(tx_qty, qty_tons > 0),
    side = "right", 
    justification = -0.2,
    binwidth = 1, 
    dotsize = 0.15) + 
   ggdist::stat_dots(
    data = subset(tx_qty, qty_tons <= 0),
    side = "right",
    justification = -0.2,
    binwidth = 1,
    dotsize = 0.15,
    color = "red") + 
  geom_hline(yintercept = 0, color = "blue") + 
  coord_flip() 
  
```

```{r}
ggplot(tx_qty, aes(x = fish_species, y = qty_tons)) + 
  geom_boxplot(
    width = .15, 
    ## remove outliers
    outlier.color = NA ## `outlier.shape = NA` or `outlier.alpha = 0` works as well
  ) +
  ## add dot plots from {ggdist} package
  ggdist::stat_dots(
    ## orientation to the right
    side = "right", 
    justification = -0.2,
    ## adjust grouping (binning) of observations 
    binwidth = 1, 
    dotsize = 0.1
  ) +
  ## add highlighted dots where qty_tons <= 0
  ggdist::stat_dots(
    data = subset(tx_qty, qty_tons <= 0),
    side = "right",
    justification = -0.2,
    binwidth = 1,
    dotsize = 0.1,
    color = "red"
  ) +
  ## add horizontal line at qty_tons = 0
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Quantity of Fish Species by Weight (Tons)",
       x = "Fish Species",
       y = "Quantity (Tons)") +
  theme(axis.text.x = element_text(hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    plot.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 7),
    plot.margin = unit(c(1, 1, 1, 1), "mm"), 
    panel.spacing = unit(0.5,"lines")) + 
  facet_grid(. ~ dest)
```

Comparing the visualisations together, we map to see the cumulative quantity of fishes received by each habor

```{r}
# Attempt to map fish species to the respective ports 


```

#### 3.3 Understanding the network of companies and ports visited (Harbor report) / region (Tping)

-   Understanding distribution of vessels in relation to paths taken

```{r}
# TBC
```

#### 3.4 Understanding seasonality of cargo at each port

-   Heatmap of cargo quantity at each port

-   Heatmap on frequency of vessel arrival at each port

```{r}
# Goal: Interactive heat map
# Zoom into specific date range, see the cargo_id for the contributing ports

#plotting heatmap
#step 1: calculate data
qty_fish_dest_time <- tx_qty %>%
  group_by(tx_date, dest, fish_species) %>%
  summarize(total_qty_tons = sum(qty_tons, na.rm = TRUE)) %>%
  ungroup()

#step 2: reshape data for heatmap

heatmap_data <- dcast(qty_fish_dest_time, tx_date + fish_species ~ dest, value.var = "total_qty_tons", fill = 0)

# Melt data for ggplot2
melted_data <- melt(heatmap_data, id.vars = c("tx_date", "fish_species"))

# step 3:  Plot the heatmap
ggplot(melted_data, aes(x = tx_date, y = fish_species, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red", name = "Total Qty (tons)") +
  labs(title = "Heatmap of Fish Species Quantity at Each Destination Over Time",
       x = "Transaction Date",
       y = "Fish Species") +
  facet_wrap(~variable, scales = "free_y", ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

3.5 Expectation of catch

```{r}
# Box plot
ggplot(tx_qty, aes(x = dest, y = qty_tons, color = fish_species)) +
  geom_boxplot() +
  labs(title = "Boxplot of Quantity in Tons by Destination",
       x = "Destination",
       y = "Quantity (tons)",
       color = "Fish Species")
```

## 4.Addressing Mini-challenge 2 questions

### 4.1 Sub-question 1: Mapping Cargo to Vessels

Preliminary comparison of total count per port

-   Does purpose or vessel type affect visits - Port Grove only Tourism
-   Comparing records, habor report includes port grove which is not present in transaction report, hence we will exclude port grove from analysis.

```{r}
# Comparing entries per port with boxplot

## tx_qty -> Type of species per port 
cargo_count <- tx_qty %>%
  group_by(dest, fish_species) %>%
  summarize(cargo_count = n()) 

ggplot(cargo_count, aes(x = dest, y = cargo_count, fill = fish_species)) +
  geom_bar(stat = "identity") +
  labs(title = "Count of Cargo per Port", x = "Port", y = "Number of Cargo", fill = "Fish Species") +
  theme_minimal()

## Hbrpt -> Type of vessel per port 
hb_v_count <- E_Hbrpt_v %>%
  group_by(port, vessel_type) %>%
  summarize(vessel_count = n()) 

ggplot(hb_v_count, aes(x = port, y = vessel_count, fill = vessel_type)) +
  geom_bar(stat = "identity") +
  labs(title = "Count of Vessel per Port", x = "Port", y = "Number of Vessel", fill = "Vessel Type") +
  theme_minimal()

## Conclusion: Exclude City of Port Grove as it is stopping over for vessels - Research, Tour, Others and not for fishing or cargo. 

# Extracting details of vessels that stop at "City of Port Grove"
port_grove_vessel <- E_Hbrpt_v %>% filter(port == "City of Port Grove")
kable(port_grove_vessel)
```

Attempt #1 to map cargo to vessel by date

```{r}
 tx_dt <- tx_qty %>% 
  # select(-c(last_edited_by, last_edited_date, date_added)) %>% 
  arrange(dest, tx_date) %>%
  select(dest, tx_date, cargo_id, fish_species, qty_tons)

# return cargo_id
datatable(tx_dt, 
          options = list(pageLength = 10), 
          filter = "top")

# plot heat map
cargo_summary <- tx_qty %>%
  group_by(dest, tx_date) %>%
  summarize(cargo_count = n())

# reshape data
cargo_summary_reshape <- dcast(cargo_summary, tx_date ~ dest, value.var = "cargo_count", fill = 0)

# Melt the data back to long format for ggplot
cargo_summary_reshape2 <- melt(cargo_summary_reshape, id.vars = "tx_date", variable.name = "dest", value.name = "cargo_count")

# Plot the heatmap
cargo_ts <- ggplot(cargo_summary_reshape2, 
                        aes(x = tx_date, y = dest, fill = cargo_count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Number of Cargo per Date at Each Harbor",
       x = "Date",
       y = "Harbor",
       fill = "Cargo Count")

ggplotly(cargo_ts)

## Add filter for city and date

```

```{r}

Hb_v_dt <- E_Hbrpt_v %>% 
  arrange(port, arr_date) %>%
  select(port, arr_date, vessel_id, vessel_company)

datatable(Hb_v_dt, 
          options = list(pageLength = 10), 
          filter = "top")

```

```{r}
vessel_summary <- E_Hbrpt_v %>%
  group_by(port, arr_date) %>%
  summarize(vessel_count = n())

# reshape data
vessel_summary_reshape <- dcast(vessel_summary, arr_date ~ port, value.var = "vessel_count", fill = 0)

# Check the reshaped data
head(vessel_summary_reshape)

# Melt the data back to long format for ggplot
vessel_summary_reshape2 <- melt(vessel_summary_reshape, id.vars = "arr_date", variable.name = "port", value.name = "vessel_count")

# Check the melted data
head(vessel_summary_reshape2)

# Plot the heatmap
vessel_ts <- ggplot(vessel_summary_reshape2, 
                    aes(x = arr_date, y = port, fill = vessel_count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Number of Vessels per Date at Each Harbor",
       x = "Date",
       y = "Harbor",
       fill = "Vessel Count")

ggplotly(vessel_ts)

```

```{r}
# Comparing count of cargo and count of vessel at habor by date and port

vessel_type_summary <- E_Hbrpt_v %>%
  group_by(port, arr_date, vessel_type) %>%
  summarize(vessel_type_count = n())

# reclass to split count by vessel type
vessel_type_summary_reshaped <- dcast(vessel_type_summary, port + arr_date ~ vessel_type, value.var = "vessel_type_count", fill = 0)

# combine individual vessel type
vessel_summary2 <- vessel_type_summary_reshaped %>%
  rename(date = arr_date, port = port) 

# combine total count of vessel
vessel_summary <- vessel_summary %>%
  rename(date = arr_date, port = port) 

# combine cargo count
cargo_summary2 <- cargo_summary %>%
  rename(date = tx_date, port = dest) # Assuming 'dest' is the correct column name in tx_qty

# Merge the datasets by date and port
by_time <- merge(vessel_summary2, vessel_summary, by = c("port", "date"), all = TRUE)
by_time <- merge(by_time, cargo_summary2, by = c("port", "date"), all = TRUE)


# test
by_time <- by_time %>% mutate(diff = vessel_count - cargo_count)

view(by_time)


```

```{r}
# Network analysis to see mapping of fish species delivered to which port at what quantity

edges <- total_qty_tons_per_dest %>%
  select(from = fish_species, to = dest, weight = total_qty_tons)

# Create graph from data frame
graph <- graph_from_data_frame(edges, directed = FALSE)

# Convert to tbl_graph object
graph_tbl <- as_tbl_graph(graph)

# Plot the network graph using ggraph
ggraph(graph_tbl, layout = 'fr') + 
  geom_edge_link(aes(width = weight), alpha = 0.8) +
  geom_node_point(aes(color = ifelse(name %in% edges$from, "Fish Species", "Destination")), size = 5) +
  geom_node_text(aes(label = name), vjust = 1.5, hjust = 1.5) +
  scale_edge_width(range = c(0.5, 5)) +
  scale_color_manual(values = c("Fish Species" = "green", "Destination" = "blue")) +
  theme_void() +
  labs(title = "Network Analysis of Fish Species per Destination", color = "Node Type", edge_width = "Total Quantity (tons)")
```

[Potential illegal activities:]{.underline}

1.  Cargo_id with fish species of `piscisosseusb6d` and `piscesfoetidaae7` where both species only found in ecological reserves

Reference: <https://isss608-ay2023-haileycsy.netlify.app/take-home_ex/take-home_ex03/take-home_ex03>

#### Quantity of cargo per city

### Network Analysis

```{r}
E_Tping_c_Fishing <- E_Tping_c %>%
  left_join(N_vessel %>% select(vessel_id, vessel_type, vessel_company), by = "vessel_id") %>%
  filter(vessel_type == "Fishing")

edges <- data.frame(
  from = E_Tping_c_Fishing$vessel_company,
  to = E_Tping_c_Fishing$ping_source
)

# Create a graph object
graph <- graph_from_data_frame(edges, directed = FALSE)

# Plot the network
plot(graph, vertex.size=10, vertex.label.cex=0.8, vertex.label.color="black",
     main="Network Visualization of Boats Movement")
```

```{r}

E_Tping_c_Fishing_SS <- E_Tping_c_Fishing %>%
  filter(vessel_company == "SouthSeafood Express Corp")

edges_ss <- data.frame(
  from = E_Tping_c_Fishing_SS$vessel_company,
  to = E_Tping_c_Fishing_SS$ping_source
)

# Create a graph object
graph_ss <- graph_from_data_frame(edges_ss, directed = FALSE)


# Create a vertex attribute to distinguish between boats, SouthSeafood, and other companies
V(graph_ss)$type <- ifelse(V(graph_ss)$name == "SouthSeafood Express Corp", "southseafood", 
                        ifelse(V(graph)$name %in% E_Tping_c_Fishing_SS$vessel_id, "boat", "location"))

# Plot the network with ggraph
ggraph(graph_ss, layout = 'fr') +  # 'fr' layout (Fruchterman-Reingold) to spread nodes nicely
  geom_edge_link(aes(width = 1), edge_colour = "grey") +  # Add edges with grey color
  geom_node_point(aes(color = type), size = 5) +  # Add nodes with different colors based on type
  geom_node_text(aes(label = name), repel = TRUE, size = 3, color = "black") +  # Add labels with repulsion
  scale_color_manual(values = c("boat" = "orange", "location" = "blue", "southseafood" = "red")) +  # Define custom colors
  theme_void() +  # Use a void theme to remove background grid
  labs(title = "Network Visualization of Boats Movement")  # Add title

```
